%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Introduction %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction}
%% refer to https://intra.ece.ucr.edu/~rlake/Whitesides_writing_res_paper.pdf for tips on introduction?
%% INTRO - TOPIC

%Why is the work important?
\sloppy % use sloppy to improve linebreaks - longer words do not overflow
Artificial Intelligence (AI) -based tools continually gain prominence as regularly leveraged tools in the
daily lives of millions of people.
Today, the significance of this technology is reflected in the current AI market size that is
estimated to be \$142 billion USD, and forecasted to increase more than tenfold by 2030~\cite{statista_artificial_2023}.
% TODO include some stats, such as daily ChatGPT users?
In addition to typical AI applications such as recommendation systems or autonomous agents, generative
models are notably increasing in popularity as well, making it one of the central research topics
in the field.
One of the most widely used implementations of generative models are Large Language Models (LLMs),
the most popular example at the moment being OpenAI's ChatGPT~\cite{openai_chatgpt_2023}.
Adoption rates of generative AI among professionals are increasing rapidly, and are already at
around 30\%~\cite{statista_us_2022}.

Large Language Models are mainly implemented in the form of text generating chatbots that can
answer seemingly any question a user might pose.
Although no expert knowledge is required to formulate a prompt and interact with an LLM-based bot, it is challenging to optimize the output, since it can vary depending on the structure, wording, and composition of the user input. %TODO cite?
Any form of model input, whether it is in the form of a task or a question, is commonly
referred to as \("\)prompting\("\) the model.
Vast application possibilities and promising future developments of LLMs make an exploration of
user prompting behavior in interactions with such models particularly interesting.

%- The objectives of the work.
In this paper, we are going to explain the fundamentals and workings of LLMs and prompting,
describe related research in the realm of user - LLM exchange, and perform our own investigation of
user behavior in these interactions.
This investigation has the objective of facilitating comprehension of existing challenges users
face when dealing with Large Language Models.
Furthermore, readers will gain a better understanding of the design of effective prompts that
enhance
generated model output.

%- Background:
% Who else has done what? How? What have we done previously?
Plenty of research has been conducted in the field of user interactions with LLMs, mainly in
regard to query reformulation strategies, studies of common user errors when prompting,
different prompt formulation strategies, and LLM limitations in general.
We will highlight relevant work in a dedicated related work section. % TODO link section?
% TODO list related work + describe briefly - fill in after completion of related work section?

%- Guidance to the reader:
% What should the reader watch for in the paper?
% What are the interesting high points? What strategy did we use?
Since the main part of this paper will be complimented by an analysis of real world examples, the
reader can expect to gain a better understanding of actual user prompting behavior.
To obtain these insights, we will leverage input data mainly gathered from the website
ShareGPT~\cite{sharegpt_sharegpt_2023}, % TODO cite
which enables users to store conversations they have had with the ChatGPT model for later retrieval
or sharing them publicly.

%- Summary/conclusion:
% What should the reader expect as conclusion?
% TODO add points as paper progresses

% TODO add info about general structure / sections of paper: "The paper is organized as follows..."
The paper is organized as follows.
The introduction provides a general overview of the paper, introducing the topic and setting the context for all
subsequent parts.
The following related work section is divided into two subsections.
We first (2.1) focus on Large Language Models (LLMs) and cover general information about their
workings, training data, text generation capabilities, real-world usage, and current limitations.
The second subsection (2.2) explores user interaction with LLMs, explains the concept of
prompting and highlights various use cases as well as related research.

Section 3 begins with a subsection (3.1) that provides an introduction to the study, outlines
our research objective, and describes the methodology and steps we will take.
Subsection 3.2 focuses on the research method used, as well as the ShareGPT and
Midjourney platforms, including details about their user base, suitability for the study, since
we leverage data samples from both as input.

Subsequently, the study results are presented in 3.3.
3.3.1 lists the findings obtained from the study, organized into predefined categories. % TODO check if really done
In 3.3.2, we analyze observable trends in user behavior and data patterns, utilizing
various visualizations.

The discussion (Section 4) starts with a subsection (4.1) that provides a synthesis of the
observed behavior in a subjective evaluation.
Sub-subsection 4.1.1 delves into the reasons why users interact with LLMs the way they do,
offering reasoning and informed assumptions.
Sub-subsection 4.1.2 explores possibilities for prompt improvement based on findings from related research.

In Subsection 4.2, we provide a future outlook, divided into an introduction of the concept of Auto
-GPT as a possible future development of prompting, as well as an overview of prompt engineering
as a newly emerging discipline in the technology sector.

The final section of this paper contains a summary of the findings and associated
discussion of the data investigation and explanations how and in which way or form we could recognize
findings from related research in our own input samples.